---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=F}
rprojroot::find_rstudio_root_file()
setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

source("00 set up.R")
load("spi prediction/league_spi_pred.RData")
pacman::p_load(
  vip,
  RColorBrewer,
  glmnet
)
```

# XGBoost Model evaluation 

## Forward Model
```{r}
league_spi_pred%>%
  pluck("fit",1)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 20,
           aesthetics = list(alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("XGB Variable Importance - FW")+
  theme_minimal()

# Top 10 
league_spi_pred%>%
  pluck("fit",1)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 10,
           aesthetics = list(fill = brewer.pal(10,"RdBu"), alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("Top 10 FW VIP")+
  theme_minimal()
```

## Midfield Model
```{r}
league_spi_pred%>%
  pluck("fit",2)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 20,
           aesthetics = list(alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("XGB Variable Importance - MF")+
  theme_minimal()

# Top 10 
league_spi_pred%>%
  pluck("fit",2)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 10,
           aesthetics = list(fill = brewer.pal(10,"RdBu"), alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("Top 10 MF VIP")+
  theme_minimal()
```

## Midfield Model
```{r}
league_spi_pred%>%
  pluck("fit",3)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 20,
           aesthetics = list(alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("XGB Variable Importance - DF")+
  theme_minimal()

# Top 10 
league_spi_pred%>%
  pluck("fit",3)%>%
  extract_fit_parsnip()%>%
  vip::vip(geom = "col", 
           num_features = 10,
           aesthetics = list(fill = brewer.pal(10,"RdBu"), alpha = 0.8)) +
  scale_y_continuous(expand = c(0, 0))+
  ggtitle("Top 10 DF VIP")+
  theme_minimal()
```

# For Goalkeeper Model

## Visualising lambda coefficients ####

Log lambda curve can be seen below:

```{r}
gk_fit<-league_spi_pred%>%
  pluck("fit",4)%>%
  extract_fit_parsnip()%>%
  pluck("fit")

gk_fit%>%
  plot(label = T, xvar = "lambda")
```

Plot paired with explanation by print shows, from left to right, how the paths of the coefficient changes against the $\ell_1$-norm, as $\lambda$ increases.

The number of non-zero coefficients is the effective degrees of freedom.

1. The number of nonzero coefficients (Df)
2. The percent (of null) deviance explained (%dev)

```{r}
gk_fit%>%
  print()

gk_fit%>%
  plot(label = T, xvar = "dev")
```

This is percent deviance explained on the training data, and is a measure of complexity of the model. We see that toward the end of the path, %deviance is not changing much but the coefficients are “blowing up” a bit. This enables us focus attention on the parts of the fit that matter.

## Obtaining the model coefficients
```{r}
league_spi_pred%>%
  pluck("fit",4)%>%
    tidy()%>%
    rename(lambda=penalty) 
```
